How crisis affect the low-income and high-income economies differently

### 2024-04-16 paper/paper.qmd
Working on Model
Update Model adjustment and mordel set up

Constant: The intercept is 5,120.936 with a standard error of 926.753, significant at the p<0.01 level. This value represents the estimated GNI per capita when all other predictors are zero, which is a theoretical baseline for the model.

Model Fit and Quality
Observations: There are 53 observations in this model, providing a somewhat small but possibly adequate sample size for initial explorations in economic data.

Log Likelihood: The log likelihood value is -432.942. While in isolation this number doesn't tell much, it helps in model comparisons, particularly when comparing nested models.

Akaike Information Criterion (AIC): The AIC is 879.883. A lower AIC indicates a better model fit relative to models with higher AIC values, assuming the same data. It is used for model selection.

@@

Constant: The intercept is 7,822.731 with a substantial standard error of 16,169.640. While not statistically significant, this value indicates the baseline GNI per capita when all predictors are zero, though this scenario is mostly theoretical given the nature of the predictors.

@@

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

@@

>>>>

@@

@@

To consider the effect of the 6 economic indicators on the level on low and high income countries' GNI per capita, I used the multiple linear regression to approach my data. The formula is as follows:

\begin{align}
\text{gni\_per\_cap} &= \alpha + \beta_1 \times \text{population}_i + \beta_2 \times \text{unem\_rate}_i + \beta_3 \times \text{lf\_par\_rate}_i \\
&\quad + \beta_4 \times \text{trade\_per}_i + \beta_5 \times \text{inflation}_i, \\
\alpha &\sim \text{Normal}(0, 2.5), \\
\beta_1 &\sim \text{Normal}(0, 2.5), \\
\beta_2 &\sim \text{Normal}(0, 2.5), \\
\beta_3 &\sim \text{Normal}(0, 2.5), \\
\beta_4 &\sim \text{Normal}(0, 2.5), \\
\beta_5 &\sim \text{Normal}(0, 2.5)
\end{align}

### Model Adjustment

For a better performance and interpretation of the model, I have applied the log transformation to my model. The log transformation is to reduce the skewness in variables that have a skewed distribution, such as `gni_per_cpa`, `gdp_growth`, `population`, etc. By transforming these variables, the distribution is more symmetric, which is closer to the normal distribution. The Log transformations also help stabilize the variance across different levels of input variables, reduces homoscedasticity.
The log transformation variables are the follows:
Log of `gni_per_cap` The dependent variable's log transformation can help in normalizing the distribution and making relationships more linear.
The log of `gdp_growth` and `population` shared similar reasons. As Economic growth rates and population can vary widely, and the impact on GNI per capita might scale logarithmically rather than linearly.

```{r}
#| echo: false
#| warning: false
#| message: false
log_lowmodel <- glm(log(gni_per_cap) ~ log(gdp_growth + 1) + log(population + 1) + unem_rate + lf_par_rate + trade_per + inflation, family = gaussian(link = "identity"), data = lowincome)

stargazer(log_lowmodel, type = "text",
          title = "Summary of Low Income Model",
          align = TRUE,
          out = "model_summary.txt")
```

```{r}
#| echo: false
#| warning: false
#| message: false
log_highmodel <- glm(log(gni_per_cap) ~ log(gdp_growth + 1) + log(population + 1) + unem_rate + lf_par_rate + trade_per + inflation,family = gaussian(link = "identity"), data = highincome)

stargazer(log_highmodel, type = "text",
          title = "Summary of High Income Model",
          align = TRUE,
          out = "model_summary.txt")
```

@@

Labor Force Participation (lf_par_rate):
Low-Income: Significant negative relationship.
High-Income: Significant positive relationship.
Insight: This contrast may indicate that in low-income countries, increases in labor force participation may include low-quality, low-income jobs, whereas in high-income countries, higher participation could be linked to high-skill, high-income roles.
Trade as Percentage of GDP (trade_per):
Both: Positive impact and significant in both models.
Insight: Trade openness appears beneficial across both income groups, enhancing GNI per capita.
Other Significant Predictors
GDP Growth (gdp_growth):
Low-Income: Insignificant.
High-Income: Significant and positive.
Insight: Economic growth plays a more critical role in high-income countries, possibly reflecting more developed financial systems and infrastructure which better translate growth into GNI per capita.
Unemployment Rate (unem_rate):
Low-Income: Insignificant.
High-Income: Significant and negative.
Insight: Unemployment negatively impacts GNI per capita in high-income countries, where unemployment may lead to significant economic losses.
Inflation (inflation):
Low-Income: Insignificant.
High-Income: Significant and negative.
Insight: Inflation is a critical factor in high-income countries, perhaps due to more developed financial markets and economies where inflation can erode real income levels more noticeably.
Model Performance Metrics
Log Likelihood and AIC:
The log likelihood is notably lower (more negative) in the high-income model, reflecting the larger scale and complexity of the dataset. The AIC is also significantly higher, suggesting a less efficient model relative to the number of parameters and the sample size compared to the low-income model.
Overall Conclusion
The comparison reveals distinct economic dynamics between low- and high-income countries. Factors like labor force participation and trade are important in both, but their impacts differ by income level. Economic growth, unemployment, and inflation are more influential in high-income contexts, suggesting different policy priorities for each group. This analysis underscores the importance of tailored economic policies that consider specific national circumstances and economic structures.
